Simulation Learning in Computer Hex

### 2011-11-04 thesis/thesis.tex
Fleshing out Hivemind design section.

\subsubsection{Learning}
The learning module is responsible for driving the evolutionary search process. Currently, Hivemind uses Evolution Strategies (see Section \ref{estrats}), although previous versions experimented with Particle Swarm Optimization. As described earlier, the learning process consists of three steps, iterated once per generation: selection, mutation, and evaluation. In general, the learning module is responsible for finding a strategy (described in Section \ref{esmods}) that generates optimal results during the evaluation step. Evaluation is performed by selecting two individuals from the population and having them play a game against each other, each individual parameterized by their strategy. The individual that wins receives points. During the selection step, individuals are culled based on points received, and the process continues. To play a game, the learning module iteratively calls into the search module with the strategy, until the tracker module reports the game is over. The tracker module is also used to collect the game results.

@@

\subsubsection{Game Tracker}

@@

Learned local play is the clear winner. Interestingly, the addition of a tenuki probability lowers performance in all cases. As expected, uniform local play significantly improves upon pure random playouts. The fact that evolutionary learning can find weights that boost performance versus uniform local in the same manner that uniform local improves upon pure random shows how much further there is to go, and shows the weaknesses of purely random simulations.

>>>>

\subsubsection{Learning}
The learning module is responsible for driving the evolutionary search process. Currently, Hivemind uses Evolution Strategies (see Section \ref{estrats}), although previous versions experimented with Particle Swarm Optimization. As described earlier, the learning process consists of three steps, iterated once per generation: selection, mutation, and evaluation. In general, the learning module is responsible for finding a strategy (described in Section \ref{esmods}) that generates optimal results during the evaluation step. Evaluation is performed by selecting two individuals from the population and having them play a game against each other, each individual parameterized by their policy weights. The individual that wins receives points. During the selection step, individuals are culled based on points received, and the process continues. To play a game, the learning module calls into the search module with the current tracker state. The search is parameterized by the policy weights. Once the tracker reports the game is over, the result (win/lose) is collected for use in the next selection step.

@@

\subsubsection{Game Tracker}
ToDo: Tracker both abstracts ``game'' notion to allow many different games with same learning and seach, and by defining what a ``game'' is, limits the scope of possible games. What games are possible with Hivemind (e.g. perfect info, 2-player, regular board), and what is off-limits/difficult (multiple players? stochastic info?).

The game tracker has four primary responsibilities.
\begin{enumerate}
\item{Maintain required game state. A single function, \texttt{Play(color, vertex)}, is used by the search tree to apply a move for the given color.}
\item{Report legal moves from a given game state. The \texttt{Legal(color, vertex)} function is used to query the legality of a given move. The search tree uses this to expand a parent node, adding all legal moves as children.}
\item{Report on the status of the winner. For Hex, this is the special \texttt{Empty} player while the game is in progress, and either \black{} or \white{} once a connected path has been achieved.}
\item{Playing out the game from the current game state. The tracker is responsible for the exact semantics of the policy weights, and for using them during the playout. The speed of the playout function is also critical, as more playouts-per-second generally result in better move selection.}
\end{enumerate}
The additional \texttt{WasPlayed(color, vertex)} function provides information required for the AMAF heuristic. A few extra functions also provide for debugging by pretty-printing the tracker's internal state.

@@

Learned local play is the clear winner. Interestingly, the addition of a tenuki probability has no or little effect on performance in all cases. As expected, uniform local play significantly improves upon pure random playouts. The fact that evolutionary learning can find weights that boost performance versus uniform local in the same manner that uniform local improves upon pure random shows how much further there is to go, and shows the weaknesses of purely random simulations.
